---
title: "Programming for Data Science Assignment"
output: html_document
---

# Declaration

I confirm that this submission is my own work and complies with course requirements.

---

# Question 1 – Product Revenue Analysis

## What I Did

- Removed cancelled transactions where `InvoiceNo` begins with “C” to ensure only valid sales were included.
- Joined 'project_orders.csv' and 'project_product_info.csv' using 'StockCode' as the primary key.
- Calculated revenue for each transaction using:

  `Revenue = Quantity × UnitPrice`

- Aggregated total revenue by product (`StockCode` and `Description`).
- Ranked products to identify the top 10 and bottom 10 performers.
- Calculated the percentage contribution of the top three products to total revenue.

## Code

```{r}
source("../R/02_q1_products.R")
```

## Interpretation

The top three products contribute **5.16% of total revenue**, indicating that revenue is not highly concentrated among a small number of products.

This suggests a diversified revenue structure where sales are distributed across many items rather than being dominated by a few bestsellers. From a risk management perspective, this reduces dependency on individual products and lowers vulnerability if one product underperforms.

However, the absence of dominant products may also suggest limited product differentiation or strong competition across the catalogue. Strategic promotion of high-performing items could increase their revenue contribution while maintaining portfolio diversification.

---

# Question 2 – High-Value Customers

## What I Did

- Removed observations with missing `CustomerID`.
- Calculated total spending per customer using aggregated revenue.
- Ranked customers to identify the top five highest-value customers.
- Analysed purchasing behaviour of these customers by summarising total quantities purchased per product.
- Visualised purchasing patterns using faceted bar charts.

## Code

```{r}
source("../R/03_q2_customers.R")
```

## Interpretation

The top five customers show significant spending concentration:

- Customer 14646: $280,206.00  
- Customer 18102: $259,657.30  
- Customer 17450: $194,550.80  
- Customer 16446: $168,472.50  
- Customer 14911: $143,825.10  

The noticeable spending gap between the highest and lowest ranked customers in the top five indicates moderate revenue concentration at the customer level.

This presents both opportunity and risk. High-value customers represent strong candidates for retention strategies such as loyalty programs or personalised marketing. However, reliance on a limited number of major customers increases vulnerability if churn occurs.

The purchasing patterns suggest bulk-buying behaviour, potentially indicating business or wholesale customers. Tailored pricing strategies or long-term contracts could strengthen these relationships.

---

# Question 3 – Country-Level Market Analysis

## What I Did

- Aggregated total revenue by country.
- Ranked countries in descending order of revenue.
- Identified the highest-quantity product within each country.
- Visualised the top 10 revenue-generating countries.

## Code

```{r}
source("../R/04_q3_countries.R")
```

## Interpretation

The **United Kingdom generated $9,003,097.96**, making it the dominant revenue source.

This indicates strong geographic concentration within a single market. While this demonstrates brand strength and demand stability in the UK, it also introduces geographic risk exposure.

If revenue from other countries is significantly lower, this suggests limited international diversification. Expanding into mid-performing markets could reduce dependence on the UK and improve long-term stability.

Variation in top-selling products across countries further supports the importance of region-specific marketing and inventory strategies rather than a uniform global approach.

---

# Question 4 – Time Trend & Consistently Low-Selling Products

## What I Did

- Converted `InvoiceDate` to proper datetime format.
- Extracted Year-Month values for time-series analysis.
- Calculated total monthly revenue.
- Identified the top three revenue-generating months.
- Determined products consistently appearing in the monthly bottom 10 revenue ranking.
- Counted the number of months each product appeared in the bottom 10.

## Code

```{r}
source("../R/05_q4_time_trends.R")
```

## Interpretation

The highest revenue months were:

- December 2009 ($254,506.80)
- December 2007 ($175,057.40)
- December 2008 ($127,885.00)

The consistent dominance of December across multiple years strongly indicates seasonal demand effects, likely driven by holiday-related purchasing behaviour and end-of-year promotions.

This suggests that inventory planning and marketing investment should increase prior to Q4 to capitalise on predictable revenue spikes.

Additionally, several products consistently underperformed:

- KITTY PENCIL ERASERS (25 months in bottom 10)
- SLEEPING CAT ERASERS (22 months in bottom 10)
- SET 12 RETRO WHITE CHALK STICKS (20 months in bottom 10)

These products exhibit structural low demand rather than temporary fluctuations. Persistently weak products may increase inventory holding costs and reduce operational efficiency.

Strategic responses could include bundling, discounting, repositioning, or discontinuation if profitability remains consistently low.

Overall, the time-series analysis highlights both predictable seasonal peaks and structural underperformance among specific products.

---

# Conclusion

The analysis demonstrates diversified product revenue, moderate customer concentration, strong geographic dependence on the United Kingdom, clear seasonal demand patterns in December, and a subset of persistently low-performing products requiring strategic review.

Collectively, these insights provide actionable recommendations for inventory management, customer retention strategies, geographic expansion, and product portfolio optimisation.